How long to simulate?

### 2015-01-25 manuscript.tex
Add methods for Experiment 1

% How do sample-based agents make decisions based on samples that take a variable amount of time?
Bayesian agents must often make judgments about situations that require computing a difficult, if not intractable, posterior distribution. One popular method to handle this issue is to use a Monte Carlo approximation to the posterior, where the agent draws some number of samples and bases its decision on those samples. However, even this approach is not always feasible for agents with limited computational computational power, as it can take thousands or even millions of samples to approach a decent approximation. The situation is further complicated by the fact that samples can vary in a variety of ways: for example, some samples may provide more information than others, and some samples may vary in how \emph{long} they take to compute. Given these constraints, what is the optimal sampling behavior for a sample-based agent?

% One and done: the best thing to do is take very few samples
Recent research by \citeA{Vul:2014ba} asked how many samples an agent should take before making a decision. Surprisingly, they found that to optimize success in the \emph{long} run, it is not useful for sample-based agents to take more than a few samples. For example, suppose that there is a reward for correct answers and no penalty for incorrect answers, and that the agent can take ten samples in the time it takes to actually execute the action that reflects the decision. Then, the optimal number of samples that the agent should base their decision on is only one! While this is obviously suboptimal for each \emph{individual} decision, over many decisions, this strategy trades off between the amount of time one can actually spend taking actions versus thinking.

% SPRT
\citeA{Vul:2014ba} examined not just the case where an agent can take a fixed number of samples, but also several policies for taking a variable number of samples depending on how useful the samples are. For example, consider the case where an agent must decide between two options. If they draw samples favoring one of the options with probability 0.95, then the agent can be fairly confident about which is the better option after only a few samples. In contrast, if the probability is 0.55, then the agent would need many samples to be confident about which option is better. Of course, the agent doesn't know what the probability is: that is the point of sampling. One strategy which addresses this issue is known as the \emph{sequential probability ratio test} \cite{wald1947sequential}, or SPRT, and has been used extensively to model a variety of decision making tasks in humans \cite<e.g.>{Gold:2007fo,Ratcliff:2008ux,Bitzer:2014ea}. An agent using SPRT accumulates evidence for one option versus another, and stops sampling once a particular threshold of evidence has been reached. So, rather than optimizing for a fixed number of samples, the agent should optimize for the threshold.

% But one and done assumes that sampling cost is fixed -- what if samples take a variable amount of time?
One piece of the puzzle not addressed by \citeA{Vul:2014ba} was the case where samples themselves take a variable amount of time. This case is not unheard of: for example, recent research studying how people reason about the physical world suggests that people may rely on samples from noisy physical simulations \cite{Smith2012,Battaglia2013,Smith:2013ug,Smith:2013th,Smith:2014tx,Ullman:2014ut,Hamrick:2015}. If people are, in fact, using physical simulations, then it is likely that each simulation will take a different amount of time to run, due to the sequential nature of the simulation. For example, if the task is to predict whether a bouncing ball will hit a green or red target first \cite{Smith:2013ug}, the time it takes to run the simulation is unknown until the ball actually hits one of the targets.

However, in nearly all the research on intuitive physics, people have been compared to a simulation-based model which took a large number of samples---rather than a small number of samples which would be optimal in the long run---and those samples have been assumed to take constant time. In their appendix, \citeA{Battaglia2013}, attempted to determine the number of samples taken by people in several tasks concerning towers of building blocks. They did find a consistent story with \citeA{Vul:2014ba}: participants in their experiments seemed to use between three and seven samples per judgment. Despite this analysis, the samples were still assumed to be constant in time cost.

% Summary
In this paper, we ask the question: what SPRT threshold should sample-based agents use when making decisions, if the time cost of taking samples is unknown? We find the counterintuitive result that the uncertainty in time cost largely doesn't matter; it is nearly optimal for the agent to act using a fixed time cost set to the expected value of the time cost. To show this, we first summarize the formal model as described by \citeA{Vul:2014ba}. Then, we extend this model to handle the case where samples take a variable amount of time. Finally, we apply the SPRT model to a behavioral case study in order to understand how people treat variable-time samples.

\section{Sampling with variable time costs}

\citeA{Vul:2014ba} previously found that optimal sample-based agents should take only a few samples before making a decision, subject to the time cost of taking those samples. In this section, we first briefly summarize their formulation of the problem. Then, we describe how this formulation must be adjusted to account for samples with variable time costs.

\subsection{Background}

% briefly go over one and done math
Here, we consider the binary (or two-alternative forced choice) case, where an agent must choose one of two hypotheses, $H_0$ or $H_1$. The agent may take samples $X_i$ from a Bernoulli distribution parameterized by an unknown parameter $p$, and from these samples estimate $\hat{p}=\frac{1}{N}\sum_{i=1}^N X_i$, where $N$ is the total number of samples. Then, the decision rule which minimizes the probability of error is $\hat{H}(X_1,\ldots{},X_N)=H_0$ when $\hat{p}<0.5$ and $\hat{H}(X_1,\ldots{},X_N)=H_1$ when $\hat{p}>0.5$.

In the best possible case, the agent takes infinite samples and chooses the maximum \emph{a posteriori} (MAP) hypothesis with probability $p$. In practice, the agent cannot take infinite samples. Thus, to determine when to stop sampling (i.e., what the value of $N$ is), we use the \emph{sequential probability ratio test} \cite{wald1947sequential}. When using a SPRT strategy, the agent accumulates $Y_N=\sum_{i=1}^N 2X_i-1$, and stops either when $Y_N=T$ or $Y_N=-T$, for some threshold number of samples $T$. Independent of the number of samples actually taken, the probability of choosing $H_1$ given that $H_1$ is the MAP hypothesis is:
\begin{equation}
\Pr[\hat{H}(Y_N)=H_1\,|\,H_1]=\frac{p^T}{p^T+(1-p)^T},
\label{eq:pr-choose-h1}
\end{equation}
and so the marginal probability of correctly choosing the MAP hypothesis is:
\begin{equation}
\Pr[\mathrm{correct}]=\frac{p^{T+1}+(1-p)^{T+1}}{p^T+(1-p)^T}.
\label{eq:pr-correct}
\end{equation}
Then, the expected utility for a decision, given a threshold and the true probability, is:
\begin{equation}
\mathbb{E}[U\,|\,T,p]=\Pr[\mathrm{correct}]\cdot{}u^{(+)}+(1-\Pr[\mathrm{correct}])\cdot{}u^{(-)},
\label{eq:expected-utility}
\end{equation}
where $u^{(+)}$ is the utility for a correct answer and $u^{(-)}$ is the utility for an incorrect answer.

We are interested not just in how much utility is gained from a single decision, but how much utility is gained from \emph{each additional sample}. Under the SPRT strategy, the probability of taking $N$ samples and then choosing $H_1$ is:
\begin{equation}
\Pr[N,H_1\,|\,T,p]=\left(\frac{2^N}{2T}\right)p^{\frac{N+T}{2}}(1-p)^{\frac{N-T}{2}}\sum_{\nu=1}^{2T-1}\cos\left(\frac{\nu\pi}{2T}\right)^{N-1}\sin\left(\frac{\nu\pi}{2T}\right)\sin\left(\frac{\nu\pi}{2}\right)
\end{equation}
for $N\geq T$ and where $N$ is of the same parity as $T$ \cite[ch.~XIV, eq. 5.7]{Feller:1968ut}. Because we are dealing with binary hypotheses, the marginal probability of $N$ samples is then:
\begin{equation}
\Pr[N\,|\,T,p]=\Pr[N,H_1\,|\,T,p]+\Pr[N,H_1\,|\,T,1-p].
\label{eq:pr-n}
\end{equation}

\begin{figure*}[t]
    \begin{center}
        \includegraphics[width=\textwidth]{figures/thresholds.pdf}
        \caption{\textbf{Rates of return and optimal thresholds.} In all plots, each line corresponds to a particular cost ratio, $t_a/t_s$. Black dots in the left two subplots show the optimal threshold $T^*$ and corresponding rate of return for each line. Left subplot: rates of return for a fixed time cost. Middle subplot: rates of return for a variable time cost. The distribution of $t_s$ was calculated empirically from simulations of a ball bouncing in a box (see Section \ref{sec:model}). Right subplot: difference in rates of return between the variable time cost and fixed time cost, approximated with Equation \ref{eq:ror-diff}.}
        \label{fig:rors}
    \end{center}
\end{figure*}

Putting together Equations \ref{eq:expected-utility} and \ref{eq:pr-n}, the expected rate or return, or utility over time, is:
\begin{equation}
\mathbb{E}[U/t\,|\,T,t_a,t_s]=\int_0^1 \sum_{N=T}^\infty \frac{\mathbb{E}[U\,|\,T,p]\cdot{}t_a}{Nt_s+t_a}\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p,
\label{eq:ror-fixed}
\end{equation}
where $t_s$ is the time it takes to draw a single sample, and $t_a$ is the time it takes to make a decision or execute an action \cite{Vul:2014ba}. Finally, given this equation, we can compute the optimal threshold for an agent using the SPRT strategy:
\begin{equation}
T^*=\mathrm{arg}\max_T \mathbb{E}[U/t\,|\,T,t_a,t_s].
\label{eq:optimal-threshold}
\end{equation}

The rate of return and corresponding optimal thresholds for a variety of time costs, as computed by Equations \ref{eq:ror-fixed} and \ref{eq:optimal-threshold}, are shown in the left subplot of Figure \ref{fig:rors}.

\subsection{Variable time costs}

% convert one and done math to variable time costs
We now turn to the case where $t_s$ is variable and unknown. Let $t_N:=\sum_{i=1}^N t_s^{(i)}$ be the time it takes to draw $N$ samples. Then, the rate of return becomes:
\begin{equation}
\mathbb{E}[U/t\,|\,T,t_a]=\int_0^1 \sum_{N=T}^\infty \mathbb{E}[U\,|\,T,p]\cdot{}t_a\cdot{}\mathbb{E}\left[\frac{1}{t_N+t_a}\right]\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p.
\label{eq:ror}
\end{equation}

Because $t_N$ is always positive, $(t_N+t_a)^{-1}$ is convex. Using Jensen's inequality:
\begin{equation}
\mathbb{E}\left[\frac{1}{t_N+t_a}\right]\geq \frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a},
\end{equation}
where the right hand side of the inequality corresponds to the original formulation by \citeA{Vul:2014ba}. Thus, the rate of return when the time cost is unknown will always be greater than when it is known. Intuitively, this makes sense: when an agent knows less to begin with, the samples it takes will give it more information, ultimately resulting in a higher rate of return.

We can approximate the extent to which samples are more useful when the time cost is unknown by computing using a Taylor expansion around $N\cdot{}\mathbb{E}[t_s]$:
\begin{equation}
\mathbb{E}\left[\frac{1}{t_N+t_a}\right]\approx \frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a}+\frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}-\ldots{}
\end{equation}
Thus, the difference is approximated by:
\begin{equation}
\mathbb{E}\left[\frac{1}{t_s+t_a}\right]-\frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a}\approx \frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}
\end{equation}
Plugging this back into Equation \ref{eq:ror}, we obtain the overall difference on the rate of return:
\begin{equation}
\Delta=\int_0^1 \sum_{N=T}^\infty \mathbb{E}[U\,|\,T,p]\cdot{}t_a\cdot{}\frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p.
\label{eq:ror-diff}
\end{equation}

In practice, this difference is quite small. The middle subplot of Figure \ref{fig:rors} shows the rates of return using a variable time cost,\footnote{The times used for this plot were computed empirically using physical simulations of a ball boucing in a box. See Section \ref{sec:model} for details.} and the right subplot shows the difference computed from \ref{eq:ror-diff}. Even the largest difference, which occurs at $T=1$ and $t_a=\mathbb{E}[t_s]$, is so small that it does not change the value of $T^*$; the difference would need to be about 100 times greater in order to have an effect. Using different distributions for $\Pr[t_s]$ may have larger effects, but even these are not substantial. For example, using a uniform distribution can change $T^*$ from 0 to 1 for smaller cost ratios.

To summarize, we find that the introduction of uncertainty into the time cost has very little effect. The implication of this result is that even if it \emph{does} takes a variable amount of time to draw a sample, an agent need not bother taking its uncertainty into account: it can rely solely on its estimate of the average sample time to determine an optimal threshold.

\section{Case study: intuitive physics}

How well does the optimal SPRT strategy actually match up with human responses on a task with variable sample costs? To answer this question, we considered the scenario where participants watch a ball bouncing around in a box, and must determine whether the ball will go through a hole in the wall.

\subsection{Model}
\label{sec:model}

@@

\subsection{Methods}

\subsubsection{Participants}

@@

\subsubsection{Stimuli}

% include figure of stimulus
The stimuli consisted of animations depicting a blue ball with a radius of 10px bouncing around in a box with dimensions 900px $\times$ 650px. All stimuli consisted of two separate animations. The first was he stimulus presentation, which had a duration of 0.775 seconds and depicted the ball moving in a particular direction. The second animation was the feedback, which picked up immediately where the first animation left off, and which had a duration of 1.5 seconds and depicted the ball either going into the hole or bouncing off the wall that contained the hole.\footnote{To enforce the constraint that the ball always travel the same distance during the feedback animation, the $x$-coordinate of the wall with the hole in it was allowed to vary across stimuli.} During the feedback animation, the ball could bounce on the other walls either 0, 1, or 2 times before going into the hole hitting the wall with the hole in it. In both animations, the ball had a velocity of 400px/s. Additionally, in the animations, the path that the ball had traveled so far was drawn with a faded gray line (see Figure \ref{fig:experiment}).

There were 48 different initial animations, and for each of these intial animations, there were four different types of feedback and two different hole sizes, giving a total of eight versions of each stimulus. The four feedback types were: ``far in'' (FI), where the ball went directly through the center of the hole; ``far miss'' (FM), where the ball missed the hole by a wide margin; ``close in'' (CI), where the ball just barely went through the hole; and `` close miss'' (CM), where the ball just barely missed the hole. The two hole sizes were 100px and 200px.

@@

In addition to the 48 experimental trials, there were seven instruction trials and eight control trials, which were the same for all participants. The control trials were designed to be extremely easy and were either of type ``straight hit'' (with a hole size of either 300px or 350px) or ``far miss'' (with a hole size of 100px). Thus, participants saw a total of 63 trials throughout the experiment.

@@

\subsubsection{Procedure}

The experiment was divided into two phases: the training phase, and the experimental phase. During the training phase, participants made judgments on the seven instruction trials in order to get used to the task. During the experimental phase, participants made judgements on the 48 experimental trials, presented in a random order, as well as the eight control trials, which were also shown in a random order, but interspersed with the experimental trials such that every 8th trial was a control trial.

On each trial, participants were shown the scene, including the initial position of the ball and the location of the hole. Participants were instructed to press the ``space'' key to begin the trial. Immediately upon pressing ``space'', the initial stimulus presentation began. As soon as the initial stimulus animation concluded, a gray box was drawn over the screen, occluding the ball (but not the line depicting the path it had traveled so far; this was left in as a reminder to participants of where the ball had come from). Participants were asked, ``will the ball go in the hole?'', and were instructed to press `q' to respond in the affirmative, and `p' to respond in the negative. Immediately after responding, text appeared saying ``Correct!'' or ``Incorrect.'', depending on the participants' response. Additionally, the gray occluder was removed, and participants were shown the feedback animation. After the feedback animation was complete, the final frame of the animation remained on the screen until participants pressed ``space'' to advance to the next trial. Figure \ref{fig:experiment} shows a few frames from one of the experimental trials.

@@

\subsubsection{Responses}

@@

\subsubsection{Response times}

@@

\subsubsection{Relationship of responses to reaction time}

@@

\subsubsection{Comparison to SPRT}

@@

In this paper, we asked the question: how should a sample-based agent act when it does not know how long it takes to draw a sample? To answer this question, we examined the optimal behavior of agents using the sequential probability ratio test (SPRT) with unknown time costs. Surprisingly, we found that the uncertainty introduced by the unknown sample times did not have a large effect: in most cases, the optimal threshold is the same regardless.

>>>>

% % How do sample-based agents make decisions based on samples that take a variable amount of time?
% Bayesian agents must often make judgments about situations that require computing a difficult, if not intractable, posterior distribution. One popular method to handle this issue is to use a Monte Carlo approximation to the posterior, where the agent draws some number of samples and bases its decision on those samples. However, even this approach is not always feasible for agents with limited computational computational power, as it can take thousands or even millions of samples to approach a decent approximation. The situation is further complicated by the fact that samples can vary in a variety of ways: for example, some samples may provide more information than others, and some samples may vary in how \emph{long} they take to compute. Given these constraints, what is the optimal sampling behavior for a sample-based agent?

% % One and done: the best thing to do is take very few samples
% Recent research by \citeA{Vul:2014ba} asked how many samples an agent should take before making a decision. Surprisingly, they found that to optimize success in the \emph{long} run, it is not useful for sample-based agents to take more than a few samples. For example, suppose that there is a reward for correct answers and no penalty for incorrect answers, and that the agent can take ten samples in the time it takes to actually execute the action that reflects the decision. Then, the optimal number of samples that the agent should base their decision on is only one! While this is obviously suboptimal for each \emph{individual} decision, over many decisions, this strategy trades off between the amount of time one can actually spend taking actions versus thinking.

% % SPRT
% \citeA{Vul:2014ba} examined not just the case where an agent can take a fixed number of samples, but also several policies for taking a variable number of samples depending on how useful the samples are. For example, consider the case where an agent must decide between two options. If they draw samples favoring one of the options with probability 0.95, then the agent can be fairly confident about which is the better option after only a few samples. In contrast, if the probability is 0.55, then the agent would need many samples to be confident about which option is better. Of course, the agent doesn't know what the probability is: that is the point of sampling. One strategy which addresses this issue is known as the \emph{sequential probability ratio test} \cite{wald1947sequential}, or SPRT, and has been used extensively to model a variety of decision making tasks in humans \cite<e.g.>{Gold:2007fo,Ratcliff:2008ux,Bitzer:2014ea}. An agent using SPRT accumulates evidence for one option versus another, and stops sampling once a particular threshold of evidence has been reached. So, rather than optimizing for a fixed number of samples, the agent should optimize for the threshold.

% % But one and done assumes that sampling cost is fixed -- what if samples take a variable amount of time?
% One piece of the puzzle not addressed by \citeA{Vul:2014ba} was the case where samples themselves take a variable amount of time. This case is not unheard of: for example, recent research studying how people reason about the physical world suggests that people may rely on samples from noisy physical simulations \cite{Smith2012,Battaglia2013,Smith:2013ug,Smith:2013th,Smith:2014tx,Ullman:2014ut,Hamrick:2015}. If people are, in fact, using physical simulations, then it is likely that each simulation will take a different amount of time to run, due to the sequential nature of the simulation. For example, if the task is to predict whether a bouncing ball will hit a green or red target first \cite{Smith:2013ug}, the time it takes to run the simulation is unknown until the ball actually hits one of the targets.

% However, in nearly all the research on intuitive physics, people have been compared to a simulation-based model which took a large number of samples---rather than a small number of samples which would be optimal in the long run---and those samples have been assumed to take constant time. In their appendix, \citeA{Battaglia2013}, attempted to determine the number of samples taken by people in several tasks concerning towers of building blocks. They did find a consistent story with \citeA{Vul:2014ba}: participants in their experiments seemed to use between three and seven samples per judgment. Despite this analysis, the samples were still assumed to be constant in time cost.

% % Summary
% In this paper, we ask the question: what SPRT threshold should sample-based agents use when making decisions, if the time cost of taking samples is unknown? We find the counterintuitive result that the uncertainty in time cost largely doesn't matter; it is nearly optimal for the agent to act using a fixed time cost set to the expected value of the time cost. To show this, we first summarize the formal model as described by \citeA{Vul:2014ba}. Then, we extend this model to handle the case where samples take a variable amount of time. Finally, we apply the SPRT model to a behavioral case study in order to understand how people treat variable-time samples.

% \section{Sampling with variable time costs}

% \citeA{Vul:2014ba} previously found that optimal sample-based agents should take only a few samples before making a decision, subject to the time cost of taking those samples. In this section, we first briefly summarize their formulation of the problem. Then, we describe how this formulation must be adjusted to account for samples with variable time costs.

% \subsection{Background}

% % briefly go over one and done math
% Here, we consider the binary (or two-alternative forced choice) case, where an agent must choose one of two hypotheses, $H_0$ or $H_1$. The agent may take samples $X_i$ from a Bernoulli distribution parameterized by an unknown parameter $p$, and from these samples estimate $\hat{p}=\frac{1}{N}\sum_{i=1}^N X_i$, where $N$ is the total number of samples. Then, the decision rule which minimizes the probability of error is $\hat{H}(X_1,\ldots{},X_N)=H_0$ when $\hat{p}<0.5$ and $\hat{H}(X_1,\ldots{},X_N)=H_1$ when $\hat{p}>0.5$.

% In the best possible case, the agent takes infinite samples and chooses the maximum \emph{a posteriori} (MAP) hypothesis with probability $p$. In practice, the agent cannot take infinite samples. Thus, to determine when to stop sampling (i.e., what the value of $N$ is), we use the \emph{sequential probability ratio test} \cite{wald1947sequential}. When using a SPRT strategy, the agent accumulates $Y_N=\sum_{i=1}^N 2X_i-1$, and stops either when $Y_N=T$ or $Y_N=-T$, for some threshold number of samples $T$. Independent of the number of samples actually taken, the probability of choosing $H_1$ given that $H_1$ is the MAP hypothesis is:
% \begin{equation}
% \Pr[\hat{H}(Y_N)=H_1\,|\,H_1]=\frac{p^T}{p^T+(1-p)^T},
% \label{eq:pr-choose-h1}
% \end{equation}
% and so the marginal probability of correctly choosing the MAP hypothesis is:
% \begin{equation}
% \Pr[\mathrm{correct}]=\frac{p^{T+1}+(1-p)^{T+1}}{p^T+(1-p)^T}.
% \label{eq:pr-correct}
% \end{equation}
% Then, the expected utility for a decision, given a threshold and the true probability, is:
% \begin{equation}
% \mathbb{E}[U\,|\,T,p]=\Pr[\mathrm{correct}]\cdot{}u^{(+)}+(1-\Pr[\mathrm{correct}])\cdot{}u^{(-)},
% \label{eq:expected-utility}
% \end{equation}
% where $u^{(+)}$ is the utility for a correct answer and $u^{(-)}$ is the utility for an incorrect answer.

% We are interested not just in how much utility is gained from a single decision, but how much utility is gained from \emph{each additional sample}. Under the SPRT strategy, the probability of taking $N$ samples and then choosing $H_1$ is:
% \begin{equation}
% \Pr[N,H_1\,|\,T,p]=\left(\frac{2^N}{2T}\right)p^{\frac{N+T}{2}}(1-p)^{\frac{N-T}{2}}\sum_{\nu=1}^{2T-1}\cos\left(\frac{\nu\pi}{2T}\right)^{N-1}\sin\left(\frac{\nu\pi}{2T}\right)\sin\left(\frac{\nu\pi}{2}\right)
% \end{equation}
% for $N\geq T$ and where $N$ is of the same parity as $T$ \cite[ch.~XIV, eq. 5.7]{Feller:1968ut}. Because we are dealing with binary hypotheses, the marginal probability of $N$ samples is then:
% \begin{equation}
% \Pr[N\,|\,T,p]=\Pr[N,H_1\,|\,T,p]+\Pr[N,H_1\,|\,T,1-p].
% \label{eq:pr-n}
% \end{equation}

% \begin{figure*}[t]
%     \begin{center}
%         \includegraphics[width=\textwidth]{figures/thresholds.pdf}
%         \caption{\textbf{Rates of return and optimal thresholds.} In all plots, each line corresponds to a particular cost ratio, $t_a/t_s$. Black dots in the left two subplots show the optimal threshold $T^*$ and corresponding rate of return for each line. Left subplot: rates of return for a fixed time cost. Middle subplot: rates of return for a variable time cost. The distribution of $t_s$ was calculated empirically from simulations of a ball bouncing in a box (see Section \ref{sec:model}). Right subplot: difference in rates of return between the variable time cost and fixed time cost, approximated with Equation \ref{eq:ror-diff}.}
%         \label{fig:rors}
%     \end{center}
% \end{figure*}

% Putting together Equations \ref{eq:expected-utility} and \ref{eq:pr-n}, the expected rate or return, or utility over time, is:
% \begin{equation}
% \mathbb{E}[U/t\,|\,T,t_a,t_s]=\int_0^1 \sum_{N=T}^\infty \frac{\mathbb{E}[U\,|\,T,p]\cdot{}t_a}{Nt_s+t_a}\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p,
% \label{eq:ror-fixed}
% \end{equation}
% where $t_s$ is the time it takes to draw a single sample, and $t_a$ is the time it takes to make a decision or execute an action \cite{Vul:2014ba}. Finally, given this equation, we can compute the optimal threshold for an agent using the SPRT strategy:
% \begin{equation}
% T^*=\mathrm{arg}\max_T \mathbb{E}[U/t\,|\,T,t_a,t_s].
% \label{eq:optimal-threshold}
% \end{equation}

% The rate of return and corresponding optimal thresholds for a variety of time costs, as computed by Equations \ref{eq:ror-fixed} and \ref{eq:optimal-threshold}, are shown in the left subplot of Figure \ref{fig:rors}.

% \subsection{Variable time costs}

% % convert one and done math to variable time costs
% We now turn to the case where $t_s$ is variable and unknown. Let $t_N:=\sum_{i=1}^N t_s^{(i)}$ be the time it takes to draw $N$ samples. Then, the rate of return becomes:
% \begin{equation}
% \mathbb{E}[U/t\,|\,T,t_a]=\int_0^1 \sum_{N=T}^\infty \mathbb{E}[U\,|\,T,p]\cdot{}t_a\cdot{}\mathbb{E}\left[\frac{1}{t_N+t_a}\right]\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p.
% \label{eq:ror}
% \end{equation}

% Because $t_N$ is always positive, $(t_N+t_a)^{-1}$ is convex. Using Jensen's inequality:
% \begin{equation}
% \mathbb{E}\left[\frac{1}{t_N+t_a}\right]\geq \frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a},
% \end{equation}
% where the right hand side of the inequality corresponds to the original formulation by \citeA{Vul:2014ba}. Thus, the rate of return when the time cost is unknown will always be greater than when it is known. Intuitively, this makes sense: when an agent knows less to begin with, the samples it takes will give it more information, ultimately resulting in a higher rate of return.

% We can approximate the extent to which samples are more useful when the time cost is unknown by computing using a Taylor expansion around $N\cdot{}\mathbb{E}[t_s]$:
% \begin{equation}
% \mathbb{E}\left[\frac{1}{t_N+t_a}\right]\approx \frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a}+\frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}-\ldots{}
% \end{equation}
% Thus, the difference is approximated by:
% \begin{equation}
% \mathbb{E}\left[\frac{1}{t_s+t_a}\right]-\frac{1}{N\cdot{}\mathbb{E}[t_s]+t_a}\approx \frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}
% \end{equation}
% Plugging this back into Equation \ref{eq:ror}, we obtain the overall difference on the rate of return:
% \begin{equation}
% \Delta=\int_0^1 \sum_{N=T}^\infty \mathbb{E}[U\,|\,T,p]\cdot{}t_a\cdot{}\frac{\mathrm{Var}(t_N)}{(N\cdot{}\mathbb{E}[t_s]+t_a)^3}\Pr[N\,|\,T,p]\Pr[p]\ \mathrm{d}p.
% \label{eq:ror-diff}
% \end{equation}

% In practice, this difference is quite small. The middle subplot of Figure \ref{fig:rors} shows the rates of return using a variable time cost,\footnote{The times used for this plot were computed empirically using physical simulations of a ball boucing in a box. See Section \ref{sec:model} for details.} and the right subplot shows the difference computed from \ref{eq:ror-diff}. Even the largest difference, which occurs at $T=1$ and $t_a=\mathbb{E}[t_s]$, is so small that it does not change the value of $T^*$; the difference would need to be about 100 times greater in order to have an effect. Using different distributions for $\Pr[t_s]$ may have larger effects, but even these are not substantial. For example, using a uniform distribution can change $T^*$ from 0 to 1 for smaller cost ratios.

% To summarize, we find that the introduction of uncertainty into the time cost has very little effect. The implication of this result is that even if it \emph{does} takes a variable amount of time to draw a sample, an agent need not bother taking its uncertainty into account: it can rely solely on its estimate of the average sample time to determine an optimal threshold.

\section{Model}
\label{sec:model}

@@

\section{Experiment 1}

Experiment 1 was a replication of the ``pong'' experiment from \cite{Smith2012}, with the major difference being a new set of stimuli and that it was run on Amazon's Mechanical Turk rather than in the lab.

\subsection{Participants}

We recruited $N=60$ participants on Amazon's Mechanical Turk using the psiTurk \cite{McDonnell12} experimental framework. Participants were treated in accordance with UC Berkeley IRB standards and were paid \$0.60 for approximately 5 minutes of work. Additionally, we excluded $N=18$ participants from analysis for failing to catch the ball on more than one control trial, leaving a total of $N=42$ participants.

\subsection{Stimuli}

There were 48 unique stimuli, each consisting of an animation depicting a blue ball with a radius of 10px bouncing around in a box with dimensions 900px $\times$ 650px, as well as a ``paddle'' that was fixed at some $x$-coordinate but moveable along the $y$-axis. Each animation had a duration of 0.775 seconds and depicted the ball moving in a particular direction with a velocity of 400px/s. Additionally, the path that the ball had traveled so far was drawn with a faded gray line (see Figure \ref{fig:experiment}).

Each animation was chosen such that were the ball to continue moving after the end of the animation, it would take exactly \TODO{} seconds to reach the $y$-coordinate of the paddle.\footnote{To enforce the constraint that the ball always travel the same distance during the feedback animation, the $x$-coordinate of the wall with the hole in it was allowed to vary across stimuli.} During this ``gap'' period in between the end of the animation and the ball reaching the $x$-coordinate of the paddle, the ball could bounce on the top, right, or bottom walls either 0, 1, or 2 times.

In addition to the 48 experimental trials, there were seven instruction trials and eight control trials. The control trials were designed to be extremely easy and consisted of a ball heading straight towards the $x$-coordinate of the paddle (with a angle of incidence of either $90^\circ$ or close to $90^\circ$. Thus, participants saw a total of 63 trials throughout the experiment.

\subsection{Procedure}

The experiment was divided into two phases: the training phase, and the experimental phase. During the training phase, participants made judgments on the seven instruction trials in order to get used to the task. During the experimental phase, participants made judgments on the 48 experimental trials, presented in a random order, as well as the eight control trials, which were also shown in a random order, but interspersed with the experimental trials such that every 8th trial was a control trial.

On each trial, participants were shown the scene, including the initial position of the ball and the location of the hole. The paddle begin at the center of the $y$-axis, and was freely moveable along the $y$-axis from the very beginning of the trial. Participants were instructed to press the ``space'' key to begin the trial. Immediately upon pressing ``space'', the initial stimulus presentation began. As soon as the initial stimulus animation concluded, a gray box was drawn over the screen, occluding the ball (but not the line depicting the path it had traveled so far; this was left in as a reminder to participants of where the ball had come from). As soon as the occluder appeared, a timer also appeared and begin counting down for \TODO{} seconds. During this time, participants had to move the paddle to try to catch the ball in the position it would be when the timer was up.

When the timer countdown finished, the paddle froze at its current location, and the occluder was removed and the full path of the ball revealed. If the ball was on the paddle, then participants were told, ``You caught the ball!''. If the ball was not on the paddle, participants were told ``Oops, you didn't catch the ball.'' Participants were then instructed to press ``space'' to continue, at which point the next trial would begin.

Throughout the experiment, participants saw two counters: one showing their total progress (the number of trials completed / total trials), and their score (the number of times they caught the ball / the number of trials completed).

\section{Experiment 2}

\subsection{Participants}

@@

\subsection{Stimuli}

The stimuli were modified versions of those used in Experiment 1, with two main differences. First, instead of a moveable paddle, there was a wall with a fixed hole in it; the wall was at the same $x$-coordinate as the paddle. Second, while the initial stimulus presentation was identical, the stimuli in Experiment 2 had a second animation which was shown after participants responded. The feedback, which picked up immediately where the first animation left off, had a duration of 1.5 seconds. It depicted the ball either going into the hole or bouncing off the wall that contained the hole.

There were 48 different initial animations (corresponding to the animations used in Experiment 1), and for each of these intial animations, there were four different types of feedback and two different hole sizes, giving a total of eight versions of each stimulus. The four feedback types were: ``far in'' (FI), where the ball went directly through the center of the hole; ``far miss'' (FM), where the ball missed the hole by a wide margin; ``close in'' (CI), where the ball just barely went through the hole; and `` close miss'' (CM), where the ball just barely missed the hole. The two hole sizes were 100px and 200px.

@@

The training trials and control trials were also the same as those used in Experiment 1, with feedback animations that were the same across participants. The feedback and hole sizes of the control trials were designed to be extremely easy and were either of type ``straight hit'' (with a hole size of either 300px or 350px) or ``far miss'' (with a hole size of 100px).

@@

\subsection{Procedure}

The procedure was very similar to that in Experiment 1, and was identical up until the point that the occluder was shown. Instead of then having a fixed amount of time to answer, participants had unlimited time. Participants were asked, ``will the ball go in the hole?'', and were instructed to press `q' to respond in the affirmative, and `p' to respond in the negative. Immediately after responding, text appeared saying ``Correct!'' or ``Incorrect.'', depending on the participants' response. Additionally, the gray occluder was removed, and participants were shown the feedback animation. After the feedback animation was complete, the final frame of the animation remained on the screen until participants pressed ``space'' to advance to the next trial. Figure \ref{fig:experiment} shows a few frames from one of the experimental trials.

\section{Results}

@@

\subsection{Responses}

@@

\subsection{Response times}

@@

\subsection{Relationship of responses to reaction time}

@@

\subsection{Comparison to SPRT}

@@

% In this paper, we asked the question: how should a sample-based agent act when it does not know how long it takes to draw a sample? To answer this question, we examined the optimal behavior of agents using the sequential probability ratio test (SPRT) with unknown time costs. Surprisingly, we found that the uncertainty introduced by the unknown sample times did not have a large effect: in most cases, the optimal threshold is the same regardless.

% We additionally ran an experiment to see how well SPRT predicts human behavior on a task which enforces unknown time costs (as well as a range of underlying probabilities). While we did find a clear relationship between people's responses and their reaction time, analagous to the type of relationship that would be expected from a SPRT-like policy. However, our SPRT model was only a mediocre explanation of people's response times---less so than even the number of times the ball bounced. However, people did not see the number of times that the ball bounced; thus, the effect of the number of bounces is a fairly clear sign that they \emph{are} using simulations in some manner. Rather than a failing of the SPRT model, this may be a failure of our simulation-based model to appropriately account for additional simulation time from bounces.

% However, based on our earlier result that unknown simulation time has little effect, even if we added extra simulation time per bounce to the model, it should not change the best-fitting threshold. Future work will investigate this claim further.
